### 2014-08-09 Docs/flowchart/flowchart.tex
more equation writing
more equation summarizing

Conservation of mass gives the same continuity equation we have with
compressible flow:
\begin{equation}
\frac{\partial \rho}{\partial t} + \nabla \cdot (\rho \Ub) = 0
\end{equation}

@@

@@

The equation of state is cast into an elliptic constraint on the
velocity field by differentiating $p_0(\rho, s, X_k)$ along particle
paths, giving:
\begin{equation}
\nabla \cdot (\beta_0 \Ub) = \beta_0 \left ( S - \frac{1}{\gammabar p_0} \frac{\partial p_0}{\partial t} \right )
\end{equation}
where
 $\beta_0$ is a density-like variable that carries background stratification, defined as
\begin{equation}
\beta_0(r,t) = \rho_0(0,t)\exp\left(\int_0^r\frac{1}{\gammabar p_0}\frac{\partial p_0}{\partial r'}dr'\right),
\end{equation}
and 
\begin{equation}
S = -\sigma\sum_k\xi_k\omegadot_k + \frac{1}{\rho p_\rho}\sum_k p_{X_k}\omegadot_k + \sigma\Hnuc + \sigma\Hext
\end{equation}
where $p_{X_k} \equiv \left. \partial p / \partial X_k
\right|_{\rho,T,X_{j,j\ne k}}$, $\xi_k \equiv \left. \partial h / 
\partial X_k \right |_{p,T,X_{j,j\ne k}}, 
p_\rho \equiv \left. \partial p/\partial \rho \right |_{T, X_k}$, and
$\sigma \equiv p_T/(\rho c_p p_\rho)$, with
$p_T \equiv \left. \partial p / \partial T \right|_{\rho, X_k}$ and
$c_p \equiv \left.  \partial h / \partial T
\right|_{p,X_k}$ is the specific heat at constant pressure.

@@

@@

We decompose the full velocity field into a base state velocity,
$w_0$, that governs the base state dynamics, and a local velocity,
$\Ubt$, that governs the local dynamics, i.e., 
\begin{equation}
\Ub = w_0(r,t)\eb_r + \Ubt(\xb,t).
\end{equation}
with
$\overline{(\Ubt\cdot\eb_r)} = 0$ and
$w_0 = \overline{(\Ub\cdot\eb_r)}$.
The velocity evolution equations are then \MarginPar{we need the $\beta_0$ here}
\begin{eqnarray}
\frac{\partial w_0}{\partial t} &=& -w_0\frac{\partial w_0}{\partial
  r} - \frac{1}{\rho_0}\frac{\partial\pi_0}{\partial r},\label{eq:w0
  evolution}\\ \frac{\partial\Ubt}{\partial t} &=& -\Ub\cdot\nabla\Ubt
  - \left(\Ubt\cdot\eb_r\right)\frac{\partial w_0}{\partial r}\eb_r -
\frac{1}{\rho}\nabla\pi +
\frac{1}{\rho_0}\frac{\partial\pi_0}{\partial r}\eb_r -
\frac{\rho-\rho_0}{\rho}g\eb_r.\label{eq:utildeupd}
\end{eqnarray}
where $\pi_0$ is the base state component of the perturbational pressure.
By laterally averaging to equation (\ref{eq:U divergence}),
we obtain a divergence constraint for $w_0$:
\begin{equation}
\nabla\cdot(\beta_0 w_0 \eb_r) = \beta_0\left(\Sbar - \frac{1}{\gammabar p_0}\frac{\partial p_0}{\partial t}\right).\label{eq:w0 divergence}
\end{equation}
The divergence constraint for $\Ubt$ can be found by subtracting
(\ref{eq:w0 divergence}) into (\ref{eq:U divergence}), resulting in
\begin{equation}
\nabla\cdot\left(\beta_0\Ubt\right) = \beta_0\left(S-\Sbar\right).\label{eq:utilde divergence}
\end{equation}

@@

Finally, we add an equation for specific enthalpy evolution to our
system.  Strictly speaking this is not necessary to close the system,
but it becomes convenient at times to define the temperature.
\begin{equation}
\frac{\partial(\rho h)}{\partial t} = -\nabla\cdot(\rho h\Ub) + \frac{Dp_0}{Dt} + \rho\Hnuc + \rho\Hext,\label{eq:enthalpy}
\end{equation}

@@

>>>>

Conservation of mass gives the same continuity equation we have with
compressible flow:
\begin{equation}
\frac{\partial \rho}{\partial t} + \nabla \cdot (\rho \Ub) = 0
\label{eq:flow:continuity}
\end{equation}

@@

The base state density evolution equation can be defined by laterally averaging the 
full continuity equation, giving:
\begin{equation}
\frac{\partial\rho_0}{\partial t} = -\nabla\cdot(\rho_0 w_0 \eb_r),\label{eq:flow:base_density}
\end{equation}

Subtracting these two yields the evolution equation for the perturbational
density, $\rho^\prime \equiv \rho - \rho_0$:
\begin{equation}
\frac{\partial\rho'}{\partial t} = -\Ub\cdot\nabla\rho' -
  \rho'\nabla\cdot\Ub - \nabla\cdot\left(\rho_0\Ubt\right) \label{eq:flow:rhoprime}
\end{equation}

As first discussed in paper III and then refined in the multilevel paper, we capture
the changes that can occur due to significant convective
overturning by imposing the constraint that $\overline{\rho'}=0$
for all time.  This gives
\begin{equation}
\frac{\partial\overline{\rho'}}{\partial t} = -\nabla\cdot(\etarho\eb_r).
\end{equation}
where
\begin{equation}
\etarho = \overline{\left(\rho'\Ub\cdot\eb_r\right)}
\end{equation}
In practice, we correct the drift by simply setting $\rho_0 =
\overline{\rho}$ after the advective update of $\rho$.  However we still need to
explicitly compute $\etarho$ since it appears in other equations.

@@

The equation of state is cast into an elliptic constraint on the
velocity field by differentiating $p_0(\rho, s, X_k)$ along particle
paths, giving:
\begin{equation}
\nabla \cdot (\beta_0 \Ub) = 
   \beta_0 \left ( S - \frac{1}{\gammabar p_0} \frac{\partial p_0}{\partial t} \right )
\label{eq:U divergence}
\end{equation}
where $\beta_0$ is a density-like variable that carries background
 stratification, defined as
\begin{equation}
\beta_0(r,t) = \rho_0(0,t)\exp\left(\int_0^r\frac{1}{\gammabar p_0}\frac{\partial p_0}{\partial r'}dr'\right),
\end{equation}
and 
\begin{equation}
S = -\sigma\sum_k\xi_k\omegadot_k + \frac{1}{\rho p_\rho}\sum_k p_{X_k}\omegadot_k + \sigma\Hnuc + \sigma\Hext
\end{equation}
where $p_{X_k} \equiv \left. \partial p / \partial X_k
\right|_{\rho,T,X_{j,j\ne k}}$, $\xi_k \equiv \left. \partial h / 
\partial X_k \right |_{p,T,X_{j,j\ne k}}, 
p_\rho \equiv \left. \partial p/\partial \rho \right |_{T, X_k}$, and
$\sigma \equiv p_T/(\rho c_p p_\rho)$, with
$p_T \equiv \left. \partial p / \partial T \right|_{\rho, X_k}$ and
$c_p \equiv \left.  \partial h / \partial T
\right|_{p,X_k}$ is the specific heat at constant pressure.

@@

In practice, we calculate $w_0$ by integrating
the one-dimensional divergence constraint:
\begin{equation}
\frac{\partial w_0}{\partial r} = \Sbar - \frac{1}{\gammabar p_0} \etarho g \enskip 
\end{equation}
Then we define
\begin{equation}
- \frac{1}{\rho_0} \frac{\partial \pizero}{\partial r} = \frac{\partial w_0}{\partial t} + 
   w_0 \frac{\partial w_0}{\partial r} \enskip, \label{eq:pizero}
\end{equation}
once $w_0$ at the old and new times is known, and the advective term is computed explicitly.
Then we can include this for completeness in the update for $\ut.$

@@

We decompose the full velocity field into a base state velocity,
$w_0$, that governs the base state dynamics, and a local velocity,
$\Ubt$, that governs the local dynamics, i.e., 
\begin{equation}
\Ub = w_0(r,t)\eb_r + \Ubt(\xb,t).
\end{equation}
with
$\overline{(\Ubt\cdot\eb_r)} = 0$ and
$w_0 = \overline{(\Ub\cdot\eb_r)}$.
The velocity evolution equations are then \MarginPar{we need the $\beta_0$ here}
\begin{eqnarray}
\frac{\partial w_0}{\partial t} &=& -w_0\frac{\partial w_0}{\partial
  r} - \frac{1}{\rho_0}\frac{\partial\pi_0}{\partial r},\label{eq:w0
  evolution}\\ \frac{\partial\Ubt}{\partial t} &=& -\Ub\cdot\nabla\Ubt
  - \left(\Ubt\cdot\eb_r\right)\frac{\partial w_0}{\partial r}\eb_r -
\frac{1}{\rho}\nabla\pi +
\frac{1}{\rho_0}\frac{\partial\pi_0}{\partial r}\eb_r -
\frac{\rho-\rho_0}{\rho}g\eb_r.\label{eq:utildeupd}
\end{eqnarray}
where $\pi_0$ is the base state component of the perturbational pressure.
By laterally averaging to equation (\ref{eq:U divergence}),
we obtain a divergence constraint for $w_0$:
\begin{equation}
\nabla\cdot(\beta_0 w_0 \eb_r) = 
    \beta_0\left(\Sbar - \frac{1}{\gammabar p_0}
           \frac{\partial p_0}{\partial t}\right).\label{eq:w0 divergence}
\end{equation}
The divergence constraint for $\Ubt$ can be found by subtracting
(\ref{eq:w0 divergence}) into (\ref{eq:U divergence}), resulting in
\begin{equation}
\nabla\cdot\left(\beta_0\Ubt\right) = \beta_0\left(S-\Sbar\right).\label{eq:utilde divergence}
\end{equation}

@@

Finally, we add an equation for specific enthalpy evolution to our
system.  Strictly speaking this is not necessary to close the system,
but it becomes convenient at times to define the temperature.
\begin{equation}
\frac{\partial(\rho h)}{\partial t} = 
   -\nabla\cdot(\rho h\Ub) + \frac{Dp_0}{Dt} + \rho\Hnuc + \rho\Hext,\label{eq:flow:enthalpy}
\end{equation}

We will often expand $Dp_0/Dt$ as
\begin{equation}
\frac{Dp_0}{Dt} = \psi + (\Ubt \cdot \er) \frac{\partial p_0}{\partial r}
\end{equation}
where we defined
\begin{equation}
\psi \equiv \frac{\partial p_0}{\partial t} + w_0 \frac{\partial p_0}{\partial r}
\end{equation}

At times, we will define a temperature equation by writing $h = h(T,p,X_k)$
and differentiating:
\begin{equation}
\label{eq:tempenthalpypsi}
\frac{DT}{Dt} = \frac{1}{\rho c_p} \left\{ \left(1 - \rho h_p\right) \left 
  [ \psi + (\Ubt \cdotb \er) \frac{\partial p_0}{\partial r} \right ]
 - \sum_k \rho \xi_k {\omegadot}_k
 + \rho \Hnuc + \rho \Hext \right \} \enskip  .
\end{equation}

@@

The base state evolution equations for density and enthalpy can be
found by averaging (\ref{eq:flow:continuity}) and (\ref{eq:flow:enthalpy})
respectively over a layer of constant radius, resulting in
\begin{eqnarray}
\frac{\partial(\rho h)_0}{\partial t} &=& -\nabla\cdot\left[(\rho h)_0w_0\eb_r\right] +
  \psi + \overline{\rho \Hnuc} + \overline{\rho \Hext}. \label{eq:flow:enthalpy_base}
\end{eqnarray}
